---
title: "Image ordering"
output: html_document
date: "2023-02-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This markdown was originally made in the dimr_bootstrap project. It's a clever way to order images by similarity to make the gif we're trying to produce a bit cleaner.

Here, we are going to implement image comparisons with Magick. Let's start with two images and see if we can get some sort of value.

```{r}
library(tidyverse)
library(here)
library(magick)
out_dir <- "im_50_cl_30"

setwd(here::here('output', out_dir))

im1 <- magick::image_read('plot1.png')
im2 <- magick::image_read('plot2.png')

magick::image_compare_dist(image = im1, reference_image = im1, metric = 'RMSE')
magick::image_compare_dist(image = im1, reference_image = im2, metric = 'RMSE')

magick::metric_types()
```
Now we're going to make a similarity matrix. Note that we make the image data much smaller. 

```{r}
setwd(here::here('output', out_dir))
files <- list.files()

ImageDistance <- function(im1, im2, metric = "MSE") {
    return(magick::image_compare_dist(image = im1, reference_image = im2, metric = 'RMSE')$distortion)
}

files <- list.files()

# Create a list to store the images, make them smaller
images <- lapply(files, function(i) {
    result <- image_read(i) %>% image_scale('10%')
})

# Create an empty matrix to store the distances
n <- length(files)
distances <- matrix(0, nrow = n, ncol = n)

# Calculate the pairwise distances between the rows of the matrix
#count <- 0
for (i in 1:n) {
  #count <- count + 1
  #print(count)
  for (j in 1:n) {
    distances[i, j] <- ImageDistance(images[[i]], images[[j]])
  }
}

rownames(distances) <- colnames(distances) <- files
```

Now let's visualize as a heatmap.

```{r}
library(pheatmap)
ph <- pheatmap::pheatmap(distances)
```
```{r}
setwd(here::here('output', out_dir))
files <- list.files()
new_order <- files[ph$tree_row$order]
#new_order <- sort(distances[1,]) %>% names()
new_order <- paste0(getwd(), "/", new_order)
```

Now we make the gif based on that new order.

```{r}
MakeGif <- function(outfile, fps = 5) {
    ## list file names and read in
    imgs <- new_order
    img_list <- lapply(imgs, function(i) {
        magick::image_read(i) %>% magick::image_scale('50%')
    })

    ## join the images together
    img_joined <- image_join(img_list)

    ## animate at 2 frames per second
    img_animated <- image_animate(img_joined, fps)

    ## save to disk
    image_write(image = img_animated,
            path = outfile)
}

MakeGif(outfile = here::here("output", out_dir, "ordered_images.gif"), fps = 10)
```



